---
title: "Compositional analysis: pairwise between three geographic locations (Zambia, South Africa, U.S)"
output: html_notebook
---

# Load in helper functions, packages, seurat object

```{r}
source("EE_vs_all_control_cohorts_analysis_helper_functions.R")
```


```{r}
library(Seurat)
library(tidyverse)
```

```{r}
disease_colors = readRDS("../reseq_analysis/disease_colors.rds")
disease_colors = c(disease_colors,"#fc9350")
saveRDS(disease_colors,"../reseq_analysis/disease_colors.rds")
```

```{r}
combined_duo = subset(combined,region=="Duodenum") %>% subset(HIV.HTLV.=="N")
```

# Composition analyses over all cells

```{r}
# Zambia vs U.S.
combined_duo_ee_no_durban = subset(combined_duo,study!="Durban")
combined_duo_ee_eoe = subset(combined_duo,study!="Resection")  %>% subset(study!="Durban")
combined_duo_ee_resection = subset(combined_duo,study!="EoE") %>%  subset(study!="Durban")
combined_duo_ee_durban = subset(combined_duo,study!="EoE") %>% subset(study!="Resection")
print(table(combined_duo_ee_durban$study))
print(table(combined_duo_ee_eoe$study))
print(table(combined_duo_ee_resection$study))

composition_ee_vs_all = fisher_composition_test(combined_duo_ee_no_durban,"tier4","study","EE",unique_types = unique(combined_duo$tier4))
composition_ee_vs_eoe = fisher_composition_test(combined_duo_ee_eoe,"tier4","study","EE",unique_types = unique(combined_duo$tier4))
composition_ee_vs_resection = fisher_composition_test(combined_duo_ee_resection,"tier4","study","EE",unique_types = unique(combined_duo$tier4))

# combining these composition results
sig_vs_all = composition_ee_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_ee_vs_eoe$sign[composition_ee_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_ee_vs_resection$sign[composition_ee_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/zambia_vs_us_all_compartments.rds")
```

```{r}
# Durban vs. U.S
combined_duo_durban_no_ee = subset(combined_duo,study!="EE")
combined_duo_durban_eoe = subset(combined_duo,study!="Resection")  %>% subset(study!="EE")
combined_duo_durban_resection = subset(combined_duo,study!="EoE") %>%  subset(study!="EE")
composition_durban_vs_all = fisher_composition_test(combined_duo_durban_no_ee,"tier4","study","Durban",unique_types = unique(combined_duo$tier4))
composition_durban_vs_eoe = fisher_composition_test(combined_duo_durban_eoe,"tier4","study","Durban",unique_types = unique(combined_duo$tier4))
composition_durban_vs_resection = fisher_composition_test(combined_duo_durban_resection,"tier4","study","Durban",unique_types = unique(combined_duo$tier4))

sig_vs_all = composition_durban_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_durban_vs_eoe$sign[composition_durban_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_durban_vs_resection$sign[composition_durban_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/durban_vs_us_all_compartments.rds")
```

```{r}
# Durban vs Zambia
combined_duo_durban_ee = subset(combined_duo,study!="EoE") %>% subset(study!="Resection")
composition_durban_vs_ee = fisher_composition_test(combined_duo_durban_ee,"tier4","study","EE",unique_types = unique(combined_duo$tier4))

final_sig  = composition_durban_vs_ee %>% filter(max_adj_pvals < 0.05)
saveRDS(final_sig,"composition/durban_vs_zambia_all_compartments.rds")
```


# Composition analyses over only epithelial cells

```{r}
epi = subset(combined_duo,tier1=="Epi")
```

```{r}
# Zambia vs U.S.
epi_ee_no_durban = subset(epi,study!="Durban")
epi_ee_eoe = subset(epi,study!="Resection")  %>% subset(study!="Durban")
epi_ee_resection = subset(epi,study!="EoE") %>%  subset(study!="Durban")
epi_ee_durban = subset(epi,study!="EoE") %>% subset(study!="Resection")
print(table(epi_ee_durban$study))
print(table(epi_ee_eoe$study))
print(table(epi_ee_resection$study))

composition_ee_vs_all = fisher_composition_test(epi_ee_no_durban,"tier4","study","EE",unique_types = unique(epi$tier4))
composition_ee_vs_eoe = fisher_composition_test(epi_ee_eoe,"tier4","study","EE",unique_types = unique(epi$tier4))
composition_ee_vs_resection = fisher_composition_test(epi_ee_resection,"tier4","study","EE",unique_types = unique(epi$tier4))

# combining these composition results
sig_vs_all = composition_ee_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_ee_vs_eoe$sign[composition_ee_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_ee_vs_resection$sign[composition_ee_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/zambia_vs_us_epi_compartments.rds")
```

```{r}
# Durban vs. U.S
epi_durban_no_ee = subset(epi,study!="EE")
epi_durban_eoe = subset(epi,study!="Resection")  %>% subset(study!="EE")
epi_durban_resection = subset(epi,study!="EoE") %>%  subset(study!="EE")
composition_durban_vs_all = fisher_composition_test(epi_durban_no_ee,"tier4","study","Durban",unique_types = unique(epi$tier4))
composition_durban_vs_eoe = fisher_composition_test(epi_durban_eoe,"tier4","study","Durban",unique_types = unique(epi$tier4))
composition_durban_vs_resection = fisher_composition_test(epi_durban_resection,"tier4","study","Durban",unique_types = unique(epi$tier4))

sig_vs_all = composition_durban_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_durban_vs_eoe$sign[composition_durban_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_durban_vs_resection$sign[composition_durban_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/durban_vs_us_epi_compartments.rds")
```

```{r}
# Durban vs Zambia
epi_durban_ee = subset(epi,study!="EoE") %>% subset(study!="Resection")
composition_durban_vs_ee = fisher_composition_test(epi_durban_ee,"tier4","study","EE",unique_types = unique(epi$tier4))

final_sig  = composition_durban_vs_ee %>% filter(max_adj_pvals < 0.05)
saveRDS(final_sig,"composition/durban_vs_zambia_epi_compartments.rds")
```
# Composition analyses over only T & NK cells

```{r}
tnk = subset(combined_duo,tier1=="TNK")
```

```{r}
# Zambia vs U.S.
tnk_ee_no_durban = subset(tnk,study!="Durban")
tnk_ee_eoe = subset(tnk,study!="Resection")  %>% subset(study!="Durban")
tnk_ee_resection = subset(tnk,study!="EoE") %>%  subset(study!="Durban")
tnk_ee_durban = subset(tnk,study!="EoE") %>% subset(study!="Resection")
print(table(tnk_ee_durban$study))
print(table(tnk_ee_eoe$study))
print(table(tnk_ee_resection$study))

composition_ee_vs_all = fisher_composition_test(tnk_ee_no_durban,"tier4","study","EE",unique_types = unique(tnk$tier4))
composition_ee_vs_eoe = fisher_composition_test(tnk_ee_eoe,"tier4","study","EE",unique_types = unique(tnk$tier4))
composition_ee_vs_resection = fisher_composition_test(tnk_ee_resection,"tier4","study","EE",unique_types = unique(tnk$tier4))

# combining these composition results
sig_vs_all = composition_ee_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_ee_vs_eoe$sign[composition_ee_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_ee_vs_resection$sign[composition_ee_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/zambia_vs_us_tnk_compartments.rds")
```

```{r}
# Durban vs. U.S
tnk_durban_no_ee = subset(tnk,study!="EE")
tnk_durban_eoe = subset(tnk,study!="Resection")  %>% subset(study!="EE")
tnk_durban_resection = subset(tnk,study!="EoE") %>%  subset(study!="EE")
composition_durban_vs_all = fisher_composition_test(tnk_durban_no_ee,"tier4","study","Durban",unique_types = unique(tnk$tier4))
composition_durban_vs_eoe = fisher_composition_test(tnk_durban_eoe,"tier4","study","Durban",unique_types = unique(tnk$tier4))
composition_durban_vs_resection = fisher_composition_test(tnk_durban_resection,"tier4","study","Durban",unique_types = unique(tnk$tier4))

sig_vs_all = composition_durban_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_durban_vs_eoe$sign[composition_durban_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_durban_vs_resection$sign[composition_durban_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/durban_vs_us_tnk_compartments.rds")
```

```{r}
# Durban vs Zambia
tnk_durban_ee = subset(tnk,study!="EoE") %>% subset(study!="Resection")
composition_durban_vs_ee = fisher_composition_test(tnk_durban_ee,"tier4","study","EE",unique_types = unique(tnk$tier4))

final_sig  = composition_durban_vs_ee %>% filter(max_adj_pvals < 0.05)
saveRDS(final_sig,"composition/durban_vs_zambia_tnk_compartments.rds")
```

# Composition analyses over only B cells, myeloid cells, and stromal cells

```{r}
bms = subset(combined_duo,tier1%in%c("B cells","Myeloid","Stromal"))
```

```{r}
# Zambia vs U.S.
bms_ee_no_durban = subset(bms,study!="Durban")
bms_ee_eoe = subset(bms,study!="Resection")  %>% subset(study!="Durban")
bms_ee_resection = subset(bms,study!="EoE") %>%  subset(study!="Durban")
bms_ee_durban = subset(bms,study!="EoE") %>% subset(study!="Resection")
print(table(bms_ee_durban$study))
print(table(bms_ee_eoe$study))
print(table(bms_ee_resection$study))

composition_ee_vs_all = fisher_composition_test(bms_ee_no_durban,"tier4","study","EE",unique_types = unique(bms$tier4))
composition_ee_vs_eoe = fisher_composition_test(bms_ee_eoe,"tier4","study","EE",unique_types = unique(bms$tier4))
composition_ee_vs_resection = fisher_composition_test(bms_ee_resection,"tier4","study","EE",unique_types = unique(bms$tier4))

# combining these composition results
sig_vs_all = composition_ee_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_ee_vs_eoe$sign[composition_ee_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_ee_vs_resection$sign[composition_ee_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/zambia_vs_us_bms_compartments.rds")
```

```{r}
# Durban vs. U.S
bms_durban_no_ee = subset(bms,study!="EE")
bms_durban_eoe = subset(bms,study!="Resection")  %>% subset(study!="EE")
bms_durban_resection = subset(bms,study!="EoE") %>%  subset(study!="EE")
composition_durban_vs_all = fisher_composition_test(bms_durban_no_ee,"tier4","study","Durban",unique_types = unique(bms$tier4))
composition_durban_vs_eoe = fisher_composition_test(bms_durban_eoe,"tier4","study","Durban",unique_types = unique(bms$tier4))
composition_durban_vs_resection = fisher_composition_test(bms_durban_resection,"tier4","study","Durban",unique_types = unique(bms$tier4))

sig_vs_all = composition_durban_vs_all %>% filter(max_adj_pvals < 0.05)
sig_vs_all$same_sign = 0
for(i in 1:nrow(sig_vs_all)){
  eoe_sign = composition_durban_vs_eoe$sign[composition_durban_vs_eoe$cell_type==sig_vs_all$cell_type[i]]
  resection_sign = composition_durban_vs_resection$sign[composition_durban_vs_resection$cell_type==sig_vs_all$cell_type[i]]
  if(eoe_sign==resection_sign & eoe_sign==sig_vs_all$sign[i]){
    sig_vs_all$same_sign[i] = 1
  }
}

final_sig = sig_vs_all %>% filter(same_sign==1)
saveRDS(final_sig,"composition/durban_vs_us_bms_compartments.rds")
```

```{r}
# Durban vs Zambia
bms_durban_ee = subset(bms,study!="EoE") %>% subset(study!="Resection")
composition_durban_vs_ee = fisher_composition_test(bms_durban_ee,"tier4","study","EE",unique_types = unique(bms$tier4))

final_sig  = composition_durban_vs_ee %>% filter(max_adj_pvals < 0.05)
saveRDS(final_sig,"composition/durban_vs_zambia_bms_compartments.rds")
```

# Plotting

## Helper functions

```{r}
plot_composition_with_labels = function(final_sig_cell_types,s_obj,use_colors){
  s_obj$cell_types = s_obj$tier4
  meta = combined_duo@meta.data %>% group_by(tier1,tier2,tier3,tier4)  %>% summarise(n())
  all_counts = generate_all_counts(s_obj@meta.data)
  temp = all_counts
  less_types = as.character(final_sig_cell_types$cell_type)
# 
#   less_types =meta$tier4[meta$tier1==compartment][meta$tier4[meta$tier1==compartment] %in% less_types]

  temp = temp[temp$cell_types %in% less_types,]
temp$cell_types <- factor(temp$cell_types,
    levels = less_types)
  options(repr.plot.width = 10, repr.plot.height = 8)
  
  temp$study = factor(temp$study,levels=c("Durban","EE","EoE","Resection"))
  temp$tier4= temp$cell_types
  print(ggplot(temp,aes(x=tier4,y=percent_of_sample,fill=study))  + geom_point(pch=21,position = position_dodge(0.75),size=3) +
   # geom_dotplot(binaxis='y', stackdir='center', dotsize=0.3,position=position_dodge(0.8)) +
    theme_classic() +
    theme(text = element_text(size=20),axis.text.x = element_text(angle = 60, hjust = 1),  panel.border = element_blank(),  
  # Remove panel grid lines
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  legend.position="top",
  # Remove panel background
  panel.background = element_blank()) +
    ggtitle("") + xlab("Cell type") + ylab("Percentage") +scale_fill_manual(values=use_colors)  + labs(fill=""))
}

plot_composition_no_labels = function(final_sig_cell_types,s_obj,use_colors){
  s_obj$cell_types = s_obj$tier4
  meta = combined_duo@meta.data %>% group_by(tier1,tier2,tier3,tier4)  %>% summarise(n())
  all_counts = generate_all_counts(s_obj@meta.data)
  temp = all_counts
  less_types = as.character(final_sig_cell_types$cell_type)
# 
#   less_types =meta$tier4[meta$tier1==compartment][meta$tier4[meta$tier1==compartment] %in% less_types]

  temp = temp[temp$cell_types %in% less_types,]
temp$cell_types <- factor(temp$cell_types,
    levels = less_types)
  options(repr.plot.width = 10, repr.plot.height = 8)
  
  temp$study = factor(temp$study,levels=c("Durban","EE","EoE","Resection"))
  temp$tier4= temp$cell_types
  print(ggplot(temp,aes(x=tier4,y=percent_of_sample,fill=study))  + geom_point(pch=21,position = position_dodge(0.75),size=3) +
   # geom_dotplot(binaxis='y', stackdir='center', dotsize=0.3,position=position_dodge(0.8)) +
    theme_classic() +
    theme(text = element_text(size=20),axis.text.x = element_blank(),  panel.border = element_blank(),  
  # Remove panel grid lines
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  legend.position="top",
  # Remove panel background
  panel.background = element_blank()) +
    ggtitle("") + xlab("Cell type") + ylab("Percentage") +scale_fill_manual(values=use_colors)  + labs(fill=""))
}

```

## Doing the plotting

### Zambia vs both U.S. cohorts


```{r}
#Epi
final_sig = readRDS("composition/ee_vs_us_epi.rds")
pdf("~/zambia_eed_figures/zambia_vs_us/epi-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,epi_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_us/epi-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,epi_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()

# TNK
final_sig = readRDS("composition/ee_vs_us_tnk.rds")
pdf("~/zambia_eed_figures/zambia_vs_us/tnk-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,tnk_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_us/tnk-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,tnk_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()

# BMS
final_sig = readRDS("composition/ee_vs_us_bms.rds")
pdf("~/zambia_eed_figures/zambia_vs_us/bms-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,bms_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_us/bms-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,bms_ee_no_ee,disease_colors[c(5,3,4)])
dev.off()
```


### Durban vs both U.S. cohorts


```{r}
#Epi
final_sig = readRDS("composition/durban_vs_us_epi.rds")
pdf("~/zambia_eed_figures/sa_vs_us/epi-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,epi_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/sa_vs_us/epi-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,epi_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()

# TNK
final_sig = readRDS("composition/durban_vs_us_tnk.rds")
pdf("~/zambia_eed_figures/sa_vs_us/tnk-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,tnk_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/sa_vs_us/tnk-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,tnk_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()

# BMS
final_sig = readRDS("composition/durban_vs_us_bms.rds")
pdf("~/zambia_eed_figures/sa_vs_us/bms-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,bms_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()
pdf("~/zambia_eed_figures/sa_vs_us/bms-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,bms_durban_no_ee,disease_colors[c(5,3,4)])
dev.off()
```


### Zambia vs Durban

```{r}
#Epi
final_sig = readRDS("composition/ee_vs_zambia_epi.rds")
pdf("~/zambia_eed_figures/zambia_vs_sa/epi-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,epi_ee_ee,disease_colors[c(1,5)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_sa/epi-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,epi_ee_ee,disease_colors[c(1,5)])
dev.off()

# TNK
final_sig = readRDS("composition/ee_vs_zambia_tnk.rds")
pdf("~/zambia_eed_figures/zambia_vs_sa/tnk-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,tnk_ee_ee,disease_colors[c(1,5)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_sa/tnk-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,tnk_ee_ee,disease_colors[c(1,5)])
dev.off()

# BMS
final_sig = readRDS("composition/ee_vs_zambia_bms.rds")
pdf("~/zambia_eed_figures/zambia_vs_sa/bms-figure/composition_labeled.pdf",useDingbats = F)
plot_composition_with_labels(final_sig,bms_ee_ee,disease_colors[c(1,5)])
dev.off()
pdf("~/zambia_eed_figures/zambia_vs_sa/bms-figure/composition_no_label.pdf",useDingbats = F)
plot_composition_no_labels(final_sig,bms_ee_ee,disease_colors[c(1,5)])
dev.off()
```

