---
title: "PROGENY and DoRothEA comparisons between EE and Durban/EoE/Resection"
output: html_notebook
---

```{r}
library(tidyverse)
library(Seurat)
library(progeny)
library(dorothea)
library(reshape2)
library(effsize)
library(viridis)

source("EE_vs_all_control_cohorts_analysis_helper_functions.R")
```

```{r}
combined = readRDS("durban_ee_eoe_resection_june152021.rds")
```

# Doing the calculations

```{r}
# combined <- progeny(combined, scale=FALSE, organism="Human", top=500, perm=1, 
#     return_assay = TRUE)
# combined<- Seurat::ScaleData(combined, assay = "progeny") 

dorothea_regulon_human <- get(data("dorothea_hs", package = "dorothea"))
## We obtain the regulons based on interactions with confidence level A, B and C
regulon <- dorothea_regulon_human %>%
    dplyr::filter(confidence %in% c("A","B","C"))
# We compute Viper Scores
combined <- run_viper(combined, regulon,
                  options = list(method = "scale", minsize = 4,
                                 eset.filter = FALSE, cores = 1,
                                 verbose = FALSE))
DefaultAssay(object = combined) <- "dorothea"
combined <- ScaleData(combined)
saveRDS(combined, "durban_ee_eoe_resection_june152021.rds")
```



# Plotting 

```{r}
combined_duo = subset(combined,region=='Duodenum') %>% subset(HIV.HTLV.=="N")
```

Looking at the epithelial results
```{r}
cell_types = c("Stem OLFM4 LGR5","Stem OLFM4","Stem cycling",
    "Ent SI","Ent APOA ALPI","BEST4",
               "Secretory precursor","Goblet","EEC","Paneth","Tuft")
subset_tiers = c(3,3,3,
          3,4,3,3,3,3,3,3)
i=1
type = cell_types[i]
# Prepping Seurat Object
if(subset_tiers[i]==4){
    sobj = subset(combined_duo,tier4==type)
}else{
    sobj = subset(combined_duo,tier3==type)
}
DefaultAssay(sobj) <- "progeny"
Idents(sobj)<-"study"
sobj_eoe = subset(sobj,study!="Resection") %>% subset(study!="Durban")
sobj_wp = subset(sobj,study!="EoE") %>% subset(study!="Durban")
sobj_sa = subset(sobj,study!="EoE") %>% subset(study!="Resection")

sobj$study[sobj$study=="EoE"]="not EE"
sobj$study[sobj$study=="Resection"]="not EE"
sobj$study[sobj$study=="Durban"]="not EE"

# Getting delta between EED & not EED scores over pathways
prog_scores = get_progeny_for_cell_type(sobj)
prog_scores$pvalue = 1
prog_scores = prog_scores %>% add_column(cell_type =rep(type,nrow(prog_scores)),.before=T)

# Calculating significant pathways
markers.vs.all = get_sobj_progeny_two_way(sobj,"study")
markers.vs.eoe = get_sobj_progeny_two_way(sobj_eoe,"study")
markers.vs.wp = get_sobj_progeny_two_way(sobj_wp,"study")
markers.vs.sa = get_sobj_progeny_two_way(sobj_sa,"study")

# need to run combine markers progeny
combined_markers = combine_markers_progeny(markers.vs.all,markers.vs.eoe,markers.vs.wp,markers.vs.sa)

for(i in 1:nrow(combined_markers)){
    prog_scores$pvalue[prog_scores$Pathway == rownames(combined_markers)[i]] = combined_markers$p_val_adj[i]
}
final_prog = prog_scores

for(i in 2:length(cell_types)){
    type = cell_types[i]
    # Prepping Seurat Object
    if(subset_tiers[i]==4){
        sobj = subset(combined_duo,tier4==type)
    }else{
        sobj = subset(combined_duo,tier3==type)
    }
    DefaultAssay(sobj) <- "progeny"
    Idents(sobj)<-"study"
    sobj_eoe = subset(sobj,study!="Resection") %>% subset(study!="Durban")
    sobj_wp = subset(sobj,study!="EoE") %>% subset(study!="Durban")
    sobj_sa = subset(sobj,study!="EoE") %>% subset(study!="Resection")
    
    sobj$study[sobj$study=="EoE"]="not EE"
    sobj$study[sobj$study=="Resection"]="not EE"
    sobj$study[sobj$study=="Durban"]="not EE"
    
    # Getting delta between EED & not EED scores over pathways
    prog_scores = get_progeny_for_cell_type(sobj)
    prog_scores$pvalue = 1
    prog_scores = prog_scores %>% add_column(cell_type =rep(type,nrow(prog_scores)),.before=T)
    
    # Calculating significant pathways
    markers.vs.all = get_sobj_progeny_two_way(sobj,"study")
    markers.vs.eoe = get_sobj_progeny_two_way(sobj_eoe,"study")
    markers.vs.wp = get_sobj_progeny_two_way(sobj_wp,"study")
    markers.vs.sa = get_sobj_progeny_two_way(sobj_sa,"study")
    
    # need to run combine markers progeny
    combined_markers = combine_markers_progeny(markers.vs.all,markers.vs.eoe,markers.vs.wp,markers.vs.sa)
    
    for(i in 1:nrow(combined_markers)){
        prog_scores$pvalue[prog_scores$Pathway == rownames(combined_markers)[i]] = combined_markers$p_val_adj[i]
    }
    
    final_prog = rbind(final_prog,prog_scores)
}
saveRDS(final_prog,"progeny_dorothea/progeny_epi.rds")
```

# ok so I am worried that there is a bug somewhere because these results should be strictly a subset of what I saw before. Unlessss things are significnat now vs the whole group weren't before, so nevermind that would make sense

```{r}
final_prog = readRDS("progeny_dorothea/progeny_epi.rds")
cell_types = c("Stem OLFM4 LGR5","Stem OLFM4","Stem cycling",
    "Ent SI","Ent APOA ALPI","BEST4",
               "Secretory precursor","Goblet","EEC","Paneth","Tuft")
final_prog_sig = final_prog %>% filter(pvalue<0.05)
final_prog_sig$minus_log_pvalue = -1*log10(final_prog_sig$pvalue)
final_prog_sig$cell_type <- factor(final_prog_sig$cell_type,levels = cell_types)

cohen_matrix = acast(final_prog, cell_type~Pathway, value.var = "cohensD")
Pathway_order = hclust(dist(t(cohen_matrix),method="euclidean"),method="ward.D")$order
final_prog$Pathway = factor(final_prog$Pathway,levels=colnames(cohen_matrix)[Pathway_order])
final_prog_sig = final_prog %>% filter(pvalue<0.05)
final_prog_sig$minus_log_pvalue = -1*log10(final_prog_sig$pvalue)
final_prog_sig$cell_type <- factor(final_prog_sig$cell_type,levels = cell_types)
pdf("~/zambia_eed_figures/three_cohorts/epithelium_figure/progeny_dotplot2.pdf",width=5,height=4.5,useDingbats = F)
ggplot(data=final_prog_sig,aes(x=cell_type,y=Pathway,color=cohensD,size=minus_log_pvalue)) +
    geom_point() + 
    scale_color_viridis(option="magma") +
    xlab("Cell Type") + theme_light() +
    theme(axis.text.x = element_text(angle = 45,hjust=1)) +
    labs(size="-log10(adj. pvalue)", colour="Cohen's D effect size \n between  EE &  not EE \n pathway scores") +
    ggtitle("")
```

# Looking at dorothea results

Basically just want to look at the highlighed DoRothEA predictions and see what those are looking like


## Running dorothea

```{r}
cell_types = c("Stem OLFM4 LGR5","Stem OLFM4","Stem cycling",
    "Ent SI","Ent APOA ALPI","BEST4",
               "Secretory precursor","Goblet","EEC","Paneth","Tuft")
subset_tiers = c(3,3,3,
          3,4,3,3,3,3,3,3)  

i=1
type = cell_types[1]
# Prepping Seurat Object
if(subset_tiers[i]==4){
    sobj = subset(combined_duo,tier4==type)
}else{
    sobj = subset(combined_duo,tier3==type)
}
DefaultAssay(sobj) <- "dorothea"
Idents(sobj)<-"study"
sobj <- ScaleData(sobj,verbose=F)
sobj_eoe = subset(sobj,study!="Resection") %>% subset(study!='Durban')
sobj_wp = subset(sobj,study!="EoE")%>% subset(study!='Durban')
sobj_sa = subset(sobj,study!="EoE") %>% subset(study!="Resection")

sobj$study[sobj$study=="EoE"]="not EE"
sobj$study[sobj$study=="Resection"]="not EE"
sobj$study[sobj$study=="Durban"]="not EE"

# Getting delta between EE & not EE scores over tfs
doro_scores = get_dorothea_for_cell_type(sobj)
doro_scores$pvalue = 1
doro_scores = doro_scores %>% add_column(cell_type =rep(type,nrow(doro_scores)),.before=T)

# Calculating significant tfs
markers.vs.all = get_sobj_dorothea_two_way(sobj,"study")
markers.vs.eoe = get_sobj_dorothea_two_way(sobj_eoe,"study")
markers.vs.wp = get_sobj_dorothea_two_way(sobj_wp,"study")
markers.vs.sa = get_sobj_dorothea_two_way(sobj_sa,"study")
combined_markers = combine_markers_dorothea(markers.vs.all,markers.vs.eoe,markers.vs.wp,markers.vs.sa)

for(i in 1:nrow(combined_markers)){
    doro_scores$pvalue[doro_scores$TF == rownames(combined_markers)[i]] = combined_markers$p_val_adj[i]
}
final_doro = doro_scores

for(i in 2:length(cell_types)){
    type = cell_types[i]
    print(type)
    # Prepping Seurat Object
    if(subset_tiers[i]==4){
        sobj = subset(combined_duo,tier4==type)
    }else{
        sobj = subset(combined_duo,tier3==type)
    }
    DefaultAssay(sobj) <- "dorothea"
    Idents(sobj)<-"study"
    sobj <- ScaleData(sobj,verbose=F)
    sobj_eoe = subset(sobj,study!="Resection") %>% subset(study!="Durban")
    sobj_wp = subset(sobj,study!="EoE") %>% subset(study!="Durban")
    sobj_sa = subset(sobj,study!="EoE") %>% subset(study!="Resection")
    
    sobj$study[sobj$study=="EoE"]="not EE"
    sobj$study[sobj$study=="Resection"]="not EE"
    sobj$study[sobj$study=="Durban"]="not EE"

    # Getting delta between EE & not EE scores over tfs
    doro_scores = get_dorothea_for_cell_type(sobj)
    doro_scores$pvalue = 1
    doro_scores = doro_scores %>% add_column(cell_type =rep(type,nrow(doro_scores)),.before=T)

    # Calculating significant tfs
    markers.vs.all = get_sobj_dorothea_two_way(sobj,"study")
    markers.vs.eoe = get_sobj_dorothea_two_way(sobj_eoe,"study")
    markers.vs.wp = get_sobj_dorothea_two_way(sobj_wp,"study")
    combined_markers = combine_markers_dorothea(markers.vs.all,markers.vs.eoe,markers.vs.wp,markers.vs.sa)

    for(i in 1:nrow(combined_markers)){
        doro_scores$pvalue[doro_scores$TF == rownames(combined_markers)[i]] = combined_markers$p_val_adj[i]
    }
    
    final_doro = rbind(final_doro,doro_scores)
}
```

```{r}
saveRDS(final_doro,"progeny_dorothea/dorothea_epi.rds")
```


```{r}
final_doro_sig = final_doro %>% filter(pvalue<0.05)
# final_doro_sig = final_doro_sig %>% filter(cohensD > 0.1 | cohensD < -0.1)
final_doro_sig$minus_log_pvalue = -1*log10(final_doro_sig$pvalue)
final_doro_sig$cell_type <- factor(final_doro_sig$cell_type,levels = cell_types)
ggplot(data=final_doro_sig,aes(x=cell_type,y=TF,fill=cohensD)) +
    geom_tile()+ 
    scale_fill_viridis(option = "magma") +
    xlab("Cell subset") + theme_classic()+ 
    theme(axis.text.x = element_text(angle = 90,hjust=1,size=3),
         axis.text.y=element_text(size=12),
         legend.position="None") +
    labs(size="-log10(adj. pvalue)", fill="Cohen's D effect size \n between  EE &  not EE \n transcription factor activity scores")+
    ggtitle("All DoRothEA TF Predictions") + ylab("Transcription Factor") + coord_flip()
```


```{r}
final_doro_sig %>% filter(cell_type=="Tuft") %>% filter(cohensD>0)
```


```{r}
final_doro_sig_select = final_doro_sig %>% filter(TF %in% c("GATA4","GATA6","STAT1","STAT6","PAX6","IRF1","IRF9",
                                                           "MYC","TCF3","TCF4",
                                                           "SMAD3","SMAD4",
                                                           "ATF2","ATF4","ATF6"))

ggplot(data=final_doro_sig_select,aes(x=cell_type,y=TF,color=cohensD,size=minus_log_pvalue)) +
    geom_point() + 
    scale_color_viridis(option="magma",limits=c(-2,2)) +
    xlab("Cell subset") + theme_light() + ylab("Transcription Factor") +
    theme(axis.text.x = element_text(angle = 45,hjust=1),legend.position="right") +
    labs(size="-log10(adj. pvalue)", colour="Cohen's D effect\n size  between\n EE &  not EE\n pathway scores") +
    ggtitle("Highlighted DoRothEA TF Predictions") + coord_flip()
```









